Learning and comparing functional connectomes across subjects

### 2013-02-24 paper.tex
Address most of reviewer 2's points

\paragraph{FMRI-based function definition}
%
Defining regions directly from the fMRI signal brings many benefits.
First, it can capture subject-specific functional information. Second, it
adapts to the signal at hand and its limitations, such as
image distortions or vascular and movement artifacts that are isolated
in ICA-like approaches. Lastly, incorporating functional information into
regional definition will result in more homogenous regions that better represent
connectivity present at the voxel level than anatomically-defined atlases such
as AAL or Harvard-Oxford \cite{craddock2012}. The simplest approach to define task-specific
regions is to use activation maps derived from standard GLM-based
analysis in a task-driven study (see for instance \cite{poldrack2011}).
Regions are extracted by thresholding the maps, or using balls around the
activation peaks. For resting-state studies, unsupervised multivariate
analysis techniques are necessary. Clustering approaches extract
full-brain parcellations \cite{craddock2012,bellec2010,yeo2011,thirion2006}, and have been shown to segment well-known functional
structures from rest data. Alternatively, decomposition
methods, such as ICA \cite{beckmann2004}, can unmix linear
combinations of multiple effects and separate out overlapping spatial maps
that capture functional networks or confounding effects. At high model
order, ICA maps define a functional parcellation \cite{kiviniemi2009}.
Extracting regions from these maps requires additional effort as they can
display fragmented spatial features and structured background noise, but
incorporating sparsity and spatial constraints in the decomposition
techniques leads to contrasted maps that outline many different
structures \cite{varoquaux2012} (see fig.\,\ref{fig:parcellations}).

\paragraph{Optimal number of regions}
%
Defining an optimal number of regions to use for whole-brain connectivity analysis bears
careful consideration. On one hand we desire a sufficiently large number of
regions to guarantee that they are functionally homogeneous regions and adequately 
represent the connectivity information present in the data. On the other hand
too many regions will render statistical inference challenging, 
result in an explosion in computational complexity,
and interfere with the interpretability of observed connections. For functional
parcellation, cross-validation methods can be employed to estimate an optimal 
number of regions based on homogeneity, the ability to reproduce 
connectivity information present at the voxel scale, and the ability to
obtain the same parcellations from independent data 
\cite{craddock2012, blumensath2012}. In general these metrics do not result in an
obvious peak at a ``best'' number of regions, but instead offer a range over
which the number of regions can be chosen based on the needs of the analysis at
hand.

@@

To study connectivity from \emph{evoked} activity with task-driven studies it
can be beneficial to run a GLM-based first-level analysis, enforcing specificity
of the measure extracted to the task. With slow event-related designs,
task-specific functional connectivity can be captured in trial-to-trial
fluctuations in the BOLD response, estimated using a GLM analysis with one
regressor per trial \cite{grillon2012,rissman2004,mennes2010}. This approach,
known as beta-series regression, has been adapted for rapid event-related
designs, using multiple GLMs to optimize deconvolution of each trial
\cite{mumford2012}.

@@

Taking a different stance on distributed variability, the ``network-based
statistics'' approach \cite{zalesky2010} draws from the hypothesis that
if, in a second-level analysis, an effect is detected on a connection
that lies in a network of strongly connected nodes, the whole network is
likely to carry an effect. Thus, they adapt cluster-level inference to
connectivity analysis, in order to mitigate the curse of multiple
comparisons.

@@

It is tempting to favor more neurobiologically-inspired models that give
descriptions close to our knowledge of the brain basic mechanisms,
however, as George Box famously said, ``all models are wrong; some models
are useful''. Depending on the question and the data at hand, a trade-off 
should be chosen between complex models based on a bio-physical description,
and simple phenomenological models such as correlation matrices. In
particular to model interactions between a large number of regions, as in
full-brain analysis, and learn a large \emph{connectome}, simple models
are to be preferred. For more hypothesis-driven studies, such as the
analysis of the mechanisms underlying a specific task, more complex models
can be preferred, if rich data is available. Automatic choice of model is
a difficult problem, however, cross-validation (as used in
\cite{varoquaux2010c,craddock2012,strother2006}) is a useful tool. The
central principle of cross-validation is to test a model on different
data than the data used to fit the model. Models too complex
for the data available will fit noise in the data, and thus generalize
poorly. The main benefit of cross-validation is that it is a
non-parametric method which does not rely strongly on modeling
assumptions\footnote{This is to be contrasted to Bayesian model
comparison, which will give well-controlled results only if the true generative
model is in the list of models compared. \cite{friston2012} argues that,
based on the Neyman-Pearson lemma, cross-validation is less powerful than
likelihood ratio tests using the full dataset. However, it is
important to keep in mind that these approaches only test for
self-consistence, as the Neyman-Pearson lemma is established under the
hypothesis that model used to define the test is indeed the 
data-generating process
\cite{neyman1933}, while in practice it is often the case that the model
gives poor fits to the data \cite{lohmann2012}. Applying test procedures
on different data than that used to fit the model, as in
cross-validation, is much more resilient to modeling errors.}.

>>>>

\paragraph{FMRI-based function definition}
%
Defining regions directly from the fMRI signal brings many benefits.
First, it can capture subject-specific functional information. Second, it
adapts to the signal at hand and its limitations, such as
image distortions or vascular and movement artifacts that are isolated
in ICA-like approaches. Lastly, incorporating functional information into
regional definition will result in more homogenous regions that better represent
connectivity present at the voxel level than anatomically-defined atlases such
as AAL or Harvard-Oxford \cite{craddock2012}. The simplest approach to define task-specific
regions is to use activation maps derived from standard GLM-based
analysis in a task-driven study (see for instance \cite{poldrack2011}).
Regions are extracted by thresholding the maps, or using balls around the
activation peaks. For resting-state studies, unsupervised multivariate
analysis techniques are necessary. Clustering approaches extract
full-brain parcellations \cite{craddock2012,bellec2010,yeo2011,thirion2006}, and have been shown to segment well-known functional
structures from rest data. Alternatively, decomposition
methods, such as ICA \cite{beckmann2004}, can unmix linear
combinations of multiple effects and separate out partially-overlapping
spatial maps
that capture functional networks or confounding effects, as for instance
with the presence of vascular structure in functional networks. At high model
order, ICA maps define a functional parcellation \cite{kiviniemi2009}.
Extracting regions from these maps requires additional effort as they can
display fragmented spatial features and structured background noise, but
incorporating sparsity and spatial constraints in the decomposition
techniques leads to contrasted maps that outline many different
structures \cite{varoquaux2012} (see fig.\,\ref{fig:parcellations}).

\paragraph{Optimal number of regions}
%
Defining an optimal number of regions to use for whole-brain connectivity analysis bears
careful consideration. On one hand we desire a sufficiently large number of
regions to guarantee that they are functionally homogeneous regions and adequately 
represent the connectivity information present in the data. On the other hand
too many regions will render statistical inference challenging, 
result in an explosion in computational complexity,
and interfere with the interpretability of observed connections. For functional
parcellation, cross-validation methods can be employed to estimate an optimal 
number of regions based on homogeneity, the ability to reproduce 
connectivity information present at the voxel scale, and the ability to
obtain the same parcellations from independent data 
\cite{craddock2012, blumensath2012}. In general these metrics do not result in an
obvious peak at a ``best'' number of regions, but instead offer a range over
which the number of regions can be chosen based on the needs of the analysis at
hand. Finally, it is important to keep in mind that there is no
universally better parcellation and associated number of regions. From a
practical standpoint, these choices will depend on the task at hand, and
more fundamentally, a good description of brain function should cover
multiple scales. Given that it is not clear that an optimal parcellation
can be identified from the sample size of a typical study, randomized
parcellation, as used in structural connectomes \cite{zalesky2010a} or
activation mapping \cite{varoquaux2012a}, may also be considered.

@@

Studying connectivity from \emph{evoked} activity with task-driven studies 
it requires disambiguating task-specific connectivity effects from
intrinsic connectivity mediated by shared neuromodulatory/task 
inputs, anatomical pathways, \emph{etc}. In this regard, it
can be beneficial to run a GLM-based first-level analysis, enforcing specificity
of the measure extracted to the task. With slow event-related designs,
task-specific functional connectivity can be captured in trial-to-trial
fluctuations in the BOLD response, estimated using a GLM analysis with one
regressor per trial \cite{grillon2012,rissman2004,mennes2010}. This approach,
known as beta-series regression, has been adapted for rapid event-related
designs, using multiple GLMs to optimize deconvolution of each trial
\cite{mumford2012}.

@@

Taking a different stance on distributed variability, the ``network-based
statistics'' approach \cite{zalesky2010} draws from the hypothesis that
if, in a second-level analysis, an effect is detected on a connection
that lies in a network of strongly connected nodes, a large sub-network is
likely to carry an effect. Thus, they adapt cluster-level inference to
connectivity analysis, in order to mitigate the curse of multiple
comparisons.

@@

It is tempting to favor more neurobiologically-inspired models that give
descriptions close to our knowledge of the brain basic mechanisms,
however, as George Box famously said, ``all models are wrong; some models
are useful''. Depending on the question and the data at hand, a trade-off 
should be chosen between complex models based on a bio-physical description,
and simple phenomenological models such as correlation matrices. In
particular to model interactions between a large number of regions, as in
full-brain analysis, and learn a large \emph{connectome}, simple models
are to be preferred. For more hypothesis-driven studies, such as the
analysis of the mechanisms underlying a specific task, more complex models
can be preferred, if rich data is available. Automatic choice of model is
a difficult problem, however, cross-validation (as used in
\cite{varoquaux2010c,craddock2012,strother2006}) is a useful tool. The
central principle of cross-validation is to test a model on different
data than the data used to fit the model. Models too complex
for the data available will fit noise in the data, and thus generalize
poorly. The main benefit of cross-validation is that it is a
non-parametric method which does not rely strongly on modeling
assumptions\footnote{This is to be contrasted to Bayesian model
comparison, which will give well-controlled results only if the true generative
model is in the list of models compared. \cite{friston2012} argues that,
based on the Neyman-Pearson lemma, cross-validation is less powerful than
likelihood ratio tests using the full dataset. However, it is
important to keep in mind that these approaches only test for
self-consistence, as the Neyman-Pearson lemma is established under the
hypothesis that the model used to define the test is indeed the 
data-generating process
\cite{neyman1933}, while in practice it is often the case that this model
gives poor fits to the data \cite{lohmann2012}. Applying test procedures
on different data than that used to fit the model, as in
cross-validation, is much more resilient to modeling errors.}.
